---
title: "Excercise 3"
author: "Christian Ortiz"
date: "11/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Excercise 3

## The Setup

```{r}
library(tidyverse)
all_states <- map_data("state")
electionData <- read_csv("2012.csv")
congress<-read_csv("womenincongress.csv")
names(electionData)[1] <- "region"
```


## 1.

Create a new variable, repProp, that gives the proportion of house representatives who are women. Use this variable to recreate the graph below which shows the proportion of house representatives who are women in each state.
```{r}
stateData <- left_join(all_states,congress,by=c("region" = "state"))
stateData <- mutate(stateData, repProp = representatives/total)
ggplot(stateData) + 
  geom_polygon(aes(x=long, y=lat, group=group, fill = repProp), color="grey50") +
  coord_quickmap()+
  scale_fill_gradient(name="Female Representatives", low="whitesmoke",high="darkred")+
  theme_classic()+
  theme(axis.ticks.y = element_blank(),axis.text.y = element_blank(), axis.ticks.x = element_blank(),axis.text.x = element_blank())+
  labs(x="",y="",title="Women in the House")
```


## 2.

Create a graph of the 2012 election for states in the South. Use the definition of the South used by the Census Bureau (see our previous work on in the introToContinuous.html notes file). You may wish to employ the function tolower() which converts text/character objects into lowercase. Your graph should look extremely similar to the one below:
```{r}
electionData$ObamaPerc <- electionData$ObamaVotes/(electionData$ObamaVotes+electionData$RomneyVotes+electionData$JohnsonVotes+electionData$SteinVotes)
electionData$RomneyPerc <- electionData$RomneyVotes/(electionData$ObamaVotes+electionData$RomneyVotes+electionData$JohnsonVotes+electionData$SteinVotes)
electionData <- merge(all_states,electionData,by="region")
```


```{r}
south_states <- c("delaware", "florida", "georgia", "maryland", "north carolina", "south carolina", "virginia", "district of columbia", "west virginia", "alabama", "kentucky", "mississippi", "tennessee", "arkansas", "louisiana", "oklahoma", "texas")
south <- filter(electionData, region %in% south_states)
ggplot(south) +
  geom_polygon(aes(x=long, y=lat, group=group, fill=ObamaPerc), color="grey50")+
  coord_quickmap()+
  labs(x="",y="",title="2012 Election Results")+
  theme_classic()+ 
  theme(axis.ticks.y = element_blank(),axis.text.y = element_blank(), axis.ticks.x = element_blank(),axis.text.x = element_blank()) + 
  scale_fill_gradient2(name="Obama's Percenatage",low="red",mid="white",high="blue",midpoint=.5)
```






